---
title: "Statistical Inference Course Project Part 2"
author: "Ignacio Ojea"
date: "June 2, 2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1. Load the data, basic exploratory analysis, and some formatting
```{r}
set.seed(2018)
data("ToothGrowth")       # Load Data
str(ToothGrowth)           # Structure of data
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```
## 2 Provide some basic summary (and visualization)
```{r, fig.width=5,fig.height=2}
summary(ToothGrowth)           # Structure of data
library(ggplot2)              # In order to provide visualization
mean.supp <- split(ToothGrowth$len, ToothGrowth$supp) # Means by supp
sapply(mean.supp, mean)
ggplot(aes(x = supp, y = len), data = ToothGrowth) + geom_boxplot(aes(fill = supp)) + xlab("Application Method") + ylab("Tooth Length [mm]") + ggtitle("Tooth Length by Application Methods") + scale_fill_manual(name = "Application method", values = c("blue","red"), labels = c("Orange juice (OJ)", "Vitamin C (VC)"))

mean.dose <- split(ToothGrowth$len, ToothGrowth$dose) # Means by dose
sapply(mean.dose, mean)
ggplot(aes(x = dose, y = len), data = ToothGrowth) + geom_boxplot(aes(fill = dose)) + xlab("Dose [mg]") + ylab("Tooth Length [mm]") + ggtitle("Tooth Length by Dose Levels") + scale_fill_manual(name = "Dose level", values = c("red", "darkgreen", "blue"), labels = c("0.5mg", "1mg", "2mg"))
```

## 3 Test to compare tooth growth by supp and dose.
Let us start by comparing tooth growth by supplement using a t-test.
```{r}
t.test(len~supp,data=ToothGrowth)
```
Since the p-value = 0.06 > 0.05 and the confidence interval [-0.17,7.57] contains zero, we can say that supplement types seems to have no impact on Tooth growth based on this test.

Now, in order to compare tooth growth by dose, we need to look at the different pairs of dose values.
```{r}
# t-test using dose amounts 0.5 and 1.0 [a 0.5 increase in dosage]
ToothGrowth.sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,0.5))
t.test(len~dose,data=ToothGrowth.sub)

# t-test using dose amounts 1.0 and 2.0 [a 1 increase in dosage]
ToothGrowth.sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(2.0,1.0))
t.test(len~dose,data=ToothGrowth.sub)

# t-test using dose amounts 0.5 and 2.0 [a 1.5 increase in dosage]
ToothGrowth.sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(2.0,0.5))
t.test(len~dose,data=ToothGrowth.sub)
```

In all three cases:

1. The p-value was approximately zero, and < 0.05.
2. The confidence interval does not include zero. 
3. Furthermore, the higher the increase in dosage, the smaller the p-value of the test.

Based on this result we can assume that the average tooth length increases with an inceasing dose, and therefore the three null hypothesis can be rejected.

## Conclusion
Given the following assumptions:

1. The sample is representative of the population.
2. The distribution of the sample means follows the Central Limit Theorem.

By observing the t-test analysis above, we can conclude that: (a) supplement delivery method has no effect on tooth growth/length, but (b) increased dosages do result in increased tooth length.
